---
title: "Math 208 H Index Function"
Author: Lily Samuel
date: "2021-12-23"
output:
  pdf_document: default
  html_document: default
---
Task: 

Find the top 10 directors in the dataset according the Hidden Gem Index (HG-H index) defined as the number of films, H, in the dataset that they have directed which have Hidden Gem Scores that are greater than or equal to H and produce them in a table with their associated HG-H index.

Solution:

```{r}
library(tidyverse)
library(utils)
library(readr)
library(dplyr)

FlixGem <- read_csv("/Users/lilysamuel/Desktop/movie_data.csv")

FlixGem <- FlixGem %>% drop_na

FlixGem <- FlixGem %>% group_split('Series.or.Movie')

print(FlixGem)

class(FlixGem)

FlixGem_Movies <- FlixGem[[1]]

print(FlixGem_Movies)

FlixGem_Movies%>%group_by(Runtime)%>%summarise(hidden_gem_score_avg= mean('Hidden.Gem.Score'))

FlixGem_Movie <- separate_rows(FlixGem_Movies, 'Country Availability', sep=", ", convert =TRUE)

class(FlixGem_Movies)
```
First, we subset to find a tibble that gives us a list of directors in the dataset, total movies they have directed, and hidden gem score average

```{r}
grouped_movies <-FlixGem_Movies %>% group_by(Director)
print(grouped_movies)

grouped_movies %>% summarise(total_movies_directed= n(),hidden_gem_score_avg = mean('Hidden.Gem.Score'))
```

Second, we create H index function

```{r}
h_index <- function(input){
      
    sorted_input <-sort(input, decreasing = F)
    for (i in 1:length(sorted_input)){
      
      result <- length(sorted_input) - i +1
      if (result <= sorted_input[[i]]){
      return(result)
    }
    
    }
  
  return(0)
      
}

input_test_1<-(c(2,2,4,4,4,4,5)) #just a test

print(h_index(input_test_1))
```

We see this function works because there are 5 characters in the vector equal to or greater than 4, therefore the H index should be 4. There test worked.

Now, apply data to the H index function to get H index of directors that we want.

```{r}
x <- separate_rows(FlixGem[[1]],Director,sep=", ",convert =TRUE)
class(x)

the_director <- unique(x$Director)
class(the_director)

h_index_vect = c() 
number = 0

for (j in the_director){
  number <- number+1
  HG_scores <- x %>% filter(Director==j) %>% select(mean = 'Hidden.Gem.Score')
  h_index_to_use <- h_index(HG_scores[[1]])
  h_index_vect <- c(h_index_vect,h_index_to_use)
  
}

print(tibble(the_director, director_H_index = h_index_vect) %>% slice_max(h_index_vect, n=10))
```









