---
title: 'Math 308: Breast Cancer Bivariate Correspondance and Multiple Correspondance Analysis'
author: "Lily Samuel"
date: "2022-04-05"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(knitr)
library(systemfonts)
library(tidyverse)
library(ggpubr )
library(FactoMineR)
library(reshape)
library(readr)
library(kableExtra)
library(factoextra)

breast_cancer <- read_csv("/Users/lilysamuel/Desktop/breast_cancer_data.csv")
head(breast_cancer)

names(breast_cancer)<-c("RecEv","AgeGrp","Meno","Size","InvNodes", "NodeCaps","DegMal","Side","Quad","Irrad")
breast_cancer<-breast_cancer %>% filter(Quad!="?",NodeCaps!="?")
breast_cancer<-breast_cancer %>% mutate(AgeGrp=paste("AG",AgeGrp,sep=""),InvNodes=paste("IN",InvNodes,sep=""))
head(breast_cancer) %>% kable(.) %>% kable_styling()
```

Task 1: 
create a two-way contingency table of counts for this data.

Solution:
```{r}
part_one<-breast_cancer %>% select(Quad,DegMal)
head(part_one)
```

Task 2:
Create ballon plots for the observed cell proportions and the expected cell proportions. Do
you find visual evidence of depedence bewteen the quadrant and the degree of malignancy? 

Solution:
```{r}
xtabs(~Quad+DegMal,data=part_one)
```

```{r}
chisq_results<-chisq.test(xtabs(~Quad+DegMal,data=part_one))

obs<-melt(chisq_results$observed)
obs<-obs%>%mutate(DegMal=factor(DegMal))
exp<-melt(chisq_results$expected)
exp<-exp%>%mutate(DegMal=factor(DegMal))

ggballoonplot(obs, label=FALSE, show.margine=F, main="Observed")
ggballoonplot(exp, label=FALSE, show.margine=F, main="Expected")

ggballoonplot(as.data.frame(t(chisq_results$observed/sum(chisq_results$observed)))) +
 ggtitle("Observed proportions")

ggballoonplot(as.data.frame(chisq_results$expected/sum(chisq_results$observed))) +
 ggtitle("Expected proportions")

ggballoonplot(as.data.frame(t(chisq_results$stdres))) +
 ggtitle("Residuals")
```
Based on what we see here, there is not a large amount of visual evidence of dependence. 
The residuals are also small, meaning only minor departures from dependence 
would be observed when doing the correspondance analysis.

Task 3:

Conduct a correspondance anaylsis for this data, being sure to complete the following
tasks:
  - Report the table of eigenvalues for the 2 components and explain how many components you
      think are sufficient to analyze the data.
  - Generate a row and column points factor map for the first two dimensions of the correspondance
      analysis (regardless of what your answer is to the first bullet point). Give separate interpretations
      of each dimension with respect to the associations between column and row points.

Solution:

```{r}
part_one_ac <- CA(xtabs(~Quad+DegMal,data=part_one))
part_one_ac$eig %>% kable(.) %>% kable_styling()
fviz_eig(part_one_ac)
fviz_ca_biplot(part_one_ac)
part_one_ac$row
part_one_ac$col
```

We see that in general there is very little evidence of structured dependence in the table. We see that all of the
coordinate values are very small for each dimension and the inertia values for all rows and columns are very
small. The rows that primarily contribute to the first dimension are the central/left_up/right_low location rows
(with left_up being contrasted with the other two). The second dimension is mostly dominated by
left_up/right_low/right_up, but these are really small in magnitude. left_low is the only row which is better
represented by the second dimension; the other four are better represented on the first dimension.
For the columns, we see a similar pattern, with very small coordinate values and small inertias. The first
dimension mostly focuses on the highest and lower degrees of malignancy (but in the same direction), without
any contribution from themiddle. The second dimension is more of a contrast between the middle malignancy
and the lowest/highest, but again the coordinate values are still quite low.
In terms of dependence between rows and columns, the first dimension really associates right_low and central
with DegMal=1,left_up with Deg_Mal 3 and the other rows/columns are not really indicated. The second
dimension shows some structure, but the coordinate values are so low, it would really be considered noise
(especially relative to the first component. )

Tast 4:
Conduct a correspondance anaylsis for this data, being sure to complete the following
tasks:
    - Report the table of eigenvalues for the first 5 components and explain how many components
      you think are sufficient to analyze the data.
    - Generate a factor map for the first two dimensions of the correspondance analysis (regardless of
        what your answer is to the first bullet point). Give a summary of which levels of which variables
        are most strongly associated with each of the first two dimensions and how you made your decisions.

Solution:

```{r}
breast_cancer<-breast_cancer %>% mutate_all(~factor(.))
```

```{r}
part_two_mca <- MCA(breast_cancer,graph=FALSE,ncp=5)
part_two_mca$eig %>% kable(.) %>% kable_styling()
fviz_eig(part_two_mca)
fviz_mca_var(part_two_mca)
fviz_mca_var(part_two_mca,choice = "var")
```

We see from this map that the first five factors explain relatively little of the variability ( (a bit less than 30% all
combined, with a maximum of 8.6% for the first factor). Probably more than 5 would be necessary here to
explain the dependence structure. If you were trying to choose a more parsimonious model, then 4 would be a
reasonable choice due to the elbow, although you would need to be careful interpreting components that
explain so little variability.
We see that for the first dimension, the levels of the variables that contribute the most are related to the
InvNodes,NodeCaps,DegMal, Irrad, and Recurrent Events. In each case of these variables, we see contrasts
between opposite ends of the spectrum. For example, InvNodes 0-2 is contrasted with all of the rest. The
second dimension is mostly driven by the Menopause and AgeGrp levels, in particular contrasting the older
ages (all groups 50 and over) and gt40 menopausal with younger (all groups 49 and younger) and premenopausal.

Task 5:

```{r}
part_two_mca_morecp <- MCA(breast_cancer,graph=FALSE,ncp=50)
sort(round(part_two_mca_morecp$var$eta2[1,],2),decreasing=TRUE)
```

We see that the next largest eta2 value is 0.14 and it only appears on dimension 27. Therefore, the
InvNodes,NodeCaps,DegMal and Irrad seeem to be most strongly associated to recurrent events. We see that
the more serious disease values (more than 2 InvNodes, NodeCaps, more serious DegMal, Irrad=yes) are all
positive along, matching RecurrentEvents=yes, with corresponding negative values for the rest matching
RecurrentEvents=no.
